---
title: "Step 4 -- HEX data"
format: html
---

<!-- For detailed instructions, see:
     https://worldbank.github.io/devPTIpack/articles/build-pti-4-hex.html -->

::: {.callout-note}
This step requires internet access to fetch data from the WB Data
Catalog and the Space2Stats API. If you are working offline, comment
out the `04-hex-data.qmd` line in `00-master.R`.
:::

## Configuration

The `HEX_RESOLUTION` and `INCLUDE_HEX_IN_APP` switches are set in
`00-master.R` and passed in as environment variables. `HEX_RESOLUTION`
controls what H3 resolution was used to build `admin9_Hexagon` in
Step 1 — `fetch_hex_data()` reads it automatically from the hex IDs.
`INCLUDE_HEX_IN_APP` controls whether hex-level polygons are included
in the deployed app.

```{r}
# Read config passed by 00-master.R via Sys.setenv(). The defaults
# below apply when this file is rendered standalone (HEX 6 = ~36 km²).
HEX_RESOLUTION     <- as.integer(Sys.getenv("PTI_HEX_RESOLUTION", "6"))
INCLUDE_HEX_IN_APP <- as.logical(Sys.getenv("PTI_INCLUDE_HEX_IN_APP", "FALSE"))
```

## Load shapes and hex layer

```{r}
library(devPTIpack)

my_shp    <- readRDS("app-data/shapes.rds")
hex_layer <- my_shp$admin9_Hexagon
hex_ids   <- hex_layer$admin9Pcod
```

## A. Discover available variables

```{r}
# Browse the full registry catalogue
list_hex_vars()

# Check live available years for a specific variable (REST-backed
# variables only — parquet-backed ones return integer(0)).
get_available_years("nighttime_lights")
```

## B. Configure hex variables — customisation point

This is the place to add or remove hex variables. **Do not** edit the
registry YAML (`inst/hex_vars_registry.yaml`) or the output Excel
directly — list variables here and re-run Step 4.

Population is always fetched automatically (needed for weighted
aggregation in Section D). Do not list it here.

```{r}
# Run list_hex_vars() to browse available variables.
# Run get_available_years("<var>") to confirm temporal coverage.
vars <- use_hex_vars(
  "flood_exposure_15cm_1in100"
  # Uncomment to add nighttime lights (pair with the `years` argument):
  # , "nighttime_lights"
  # Uncomment to add GHSL built-up area:
  # , "builtup_area"
  # `years` applies to all temporal variables in this call:
  # , years = c(2020, 2022)
)
```

## C. Fetch hex data

```{r}
hex_data <- fetch_hex_data(hex_ids = hex_ids, vars = vars)
```

## C′. Optional: merge your own local parquet

```{r}
#| eval: false
# Uncomment and adapt if you have project-specific indicators.
# Local parquet must be long format, no population column.
#
# my_data <- arrow::open_dataset("data-raw/my_indicators.parquet") |>
#   dplyr::filter(my_hex_col %in% hex_ids) |>
#   dplyr::collect() |>
#   dplyr::rename(hex_id = my_hex_col) |>
#   tidyr::pivot_wider(
#     names_from  = year,
#     values_from = my_variable,
#     names_prefix = "my_var_"
#   )
#
# hex_data <- dplyr::full_join(hex_data, my_data, by = "hex_id")
```

## D. Aggregate to admin polygons

```{r}
aggregated <- aggregate_hex_to_shapes(
  hex_data  = hex_data,
  hex_layer = hex_layer,
  shp_dta   = my_shp,
  strategy  = list(
    # Population is always aggregated internally as weight = "none",
    # fun = "sum". Do not set it here.
    .default = c(weight = "pop", fun = "mean")
    # Per-variable overrides (uncomment if needed):
    # , nighttime_lights = c(weight = "none", fun = "mean")
  )
)
```

## E. Build metadata Excel

```{r}
build_hex_metadata(
  aggregated   = aggregated,
  shp_dta      = my_shp,
  # indicator_config needed only for non-registry variables:
  # indicator_config = tibble::tibble(
  #   canonical_name = "my_var",
  #   var_name       = "My Variable ({year})",
  #   pillar_group   = 4,
  #   pillar_name    = "Custom"
  # ),
  country_name = "Rwanda",
  output_path  = "app-data/metadata-hex.xlsx",
  include_hex  = INCLUDE_HEX_IN_APP
  # include_population = TRUE  # uncomment to add population as a visible
  #                            # map indicator (excluded by default to avoid
  #                            # collision with Step 3 population variables)
)
```

## F. Summary of fetched hex variables

```{r}
pti_summary_table(hex_data, type = "hex")
```

## Ordering note

This step is **order-flexible** — it can be run before or after
Step 3. The two outputs (`metadata-user.xlsx` and `metadata-hex.xlsx`)
are independent intermediate files merged together in Step 5.
