---
title: "{{APP_NAME}} -- Data Reference"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
  pdf:
    toc: true
    number-sections: false
execute:
  echo: false
---

<!-- For detailed instructions, see:
     https://worldbank.github.io/devPTIpack/articles/build-pti-5-compile.html -->

::: {.content-visible when-format="html"}
[⬇ Download PDF version](pti-metadata.pdf){.btn .btn-outline-primary}
:::

This is the data-quality and metadata reference for the compiled PTI
app. It is generated from `app-data/shapes.rds` and
`app-data/metadata.xlsx` -- run Step 5 (`05-compile.qmd`) first.

```{r}
#| label: setup
#| message: false
#| warning: false
library(devPTIpack)

my_shp   <- readRDS("app-data/shapes.rds")
compiled <- fct_template_reader("app-data/metadata.xlsx")

# Em dash for empty metadata cells.
or_dash <- function(x) {
  if (length(x) == 0L || is.na(x) || !nzchar(as.character(x))) {
    "—"
  } else {
    as.character(x)
  }
}
```

## Spatial boundaries

One panel per administrative level in the compiled shapes, with a
polygon-count and area summary.

```{r}
#| label: boundaries-map
#| fig-width: 8
#| fig-height: 5
pti_plot_boundaries(my_shp)
```

```{r}
#| label: boundaries-table
boundary_summary <- do.call(rbind, lapply(names(my_shp), function(nm) {
  layer    <- my_shp[[nm]]
  area_km2 <- as.numeric(sf::st_area(layer)) / 1e6
  data.frame(
    Layer               = nm,
    Polygons            = nrow(layer),
    `Mean area (km2)`   = round(mean(area_km2), 1),
    `Min area (km2)`    = round(min(area_km2), 1),
    `Max area (km2)`    = round(max(area_km2), 1),
    check.names = FALSE
  )
}))
knitr::kable(boundary_summary)
```

## Variable reference

Every indicator in `metadata.xlsx`, ordered by pillar then
`var_order`. Each shows a choropleth at its source admin level, the
value distribution with decile markers, and summary statistics.

```{r}
#| label: variable-reference
#| results: asis
#| fig-width: 7
#| fig-height: 4
#| message: false
#| warning: false
md <- compiled$metadata
md <- md[order(md$pillar_group, md$var_order), , drop = FALSE]

for (i in seq_len(nrow(md))) {
  v   <- md[i, ]
  lvl <- v$spatial_level
  vc  <- v$var_code

  cat("\n\n### ", or_dash(v$var_name), "\n\n", sep = "")
  cat(
    "**Pillar:** ", or_dash(v$pillar_name),
    "  |  **Units:** ", or_dash(v$var_units),
    "  |  **Admin level:** ", or_dash(lvl),
    "\n\n", sep = ""
  )

  layer_tbl <- compiled[[lvl]]
  geom      <- my_shp[[lvl]]

  if (is.null(layer_tbl) || is.null(geom) || !vc %in% names(layer_tbl)) {
    cat("_No data found for this variable at ", or_dash(lvl), "._\n\n",
        sep = "")
    next
  }

  # Panel data: report the most recent year only. Skip when `year` is
  # absent or entirely NA (cross-sectional data) -- filtering on an
  # all-NA column would otherwise select NA rows and blank every value.
  if ("year" %in% names(layer_tbl) && any(!is.na(layer_tbl$year))) {
    keep_year <- max(layer_tbl$year, na.rm = TRUE)
    layer_tbl <- layer_tbl[
      !is.na(layer_tbl$year) & layer_tbl$year == keep_year, ,
      drop = FALSE
    ]
  }

  # Join the indicator column onto the layer geometry for the map.
  pcod   <- paste0(sub("_.*", "", lvl), "Pcod")
  joined <- dplyr::left_join(
    geom, layer_tbl[, c(pcod, vc)], by = pcod
  )

  print(pti_plot_choropleth(joined, vc))
  print(pti_plot_histogram(layer_tbl, vc))

  vals <- layer_tbl[[vc]]
  qs   <- stats::quantile(vals, c(.1, .25, .5, .75, .9), na.rm = TRUE)
  stat_tbl <- data.frame(
    Statistic = c("Min", "Q10", "Q25", "Median", "Mean",
                  "Q75", "Q90", "Max", "SD", "NA count"),
    Value = c(
      min(vals, na.rm = TRUE), qs[[1]], qs[[2]], qs[[3]],
      mean(vals, na.rm = TRUE), qs[[4]], qs[[5]],
      max(vals, na.rm = TRUE), stats::sd(vals, na.rm = TRUE),
      sum(is.na(vals))
    ),
    check.names = FALSE
  )
  cat(knitr::kable(stat_tbl, digits = 2), sep = "\n")
  cat("\n\n")

  # Three variables per page.
  if (i %% 3 == 0 && i < nrow(md)) cat("{{< pagebreak >}}\n\n")
}
```
