---
title: "Step 5 -- Compile and finalise"
format: html
---

<!-- For detailed instructions, see:
     https://worldbank.github.io/devPTIpack/articles/build-pti-5-compile.html -->

This step takes the intermediate metadata files produced by Steps 3
(and optionally Step 4) and `compile_pti_data()` produces the three
deployment-ready artefacts under `app-data/`:

- `app-data/metadata.xlsx` -- canonical merged metadata workbook.
- `app-data/shapefiles.zip` -- GeoJSON archive (one file per admin
  level) for download from the deployed app.
- `app-data/pti-metadata.pdf` -- printable indicator reference,
  mapping each indicator at its most-disaggregated admin level (or
  at the explicit `spatial_level` you set in metadata).

`compile_pti_data()` also re-runs `validate_geometries()` and
`validate_metadata()` on the combined inputs and prints a verbose
CLI summary. On validation failure it errors by default; pass
`error_on_fail = FALSE` to inspect the structured result instead.

## Variable overrides

This table controls what each indicator does in the deployed app.
**Do not** edit the metadata Excel files (`metadata-user.xlsx` /
`metadata-hex.xlsx`) directly -- Steps 3 and 4 regenerate them and any
manual edit is overwritten. Set overrides here instead.

- `in_pti = FALSE` -- variable excluded from the composite PTI score.
- `in_explorer = FALSE` -- variable hidden from the Data Explorer tab.

`canonical_name` is the variable's `var_code` as shown in the data
summary further down. Mind the commas when uncommenting rows.

```{r}
var_overrides <- tibble::tribble(
  ~canonical_name, ~in_pti, ~in_explorer
  # , "flood_exposure_15cm_1in100", FALSE, TRUE   # keep in explorer, drop from PTI
  # , "poverty_rate",                TRUE, FALSE  # in PTI, hide from explorer
)
```

## Run

`metadata-hex.xlsx` (Step 4's output) is included automatically when
it exists -- no manual edit needed. If you skipped Step 4, only the
user metadata is compiled.

```{r}
#| message: false
library(devPTIpack)

hex_mtdt_path  <- "app-data/metadata-hex.xlsx"
metadata_paths <- c(
  "app-data/metadata-user.xlsx",
  if (file.exists(hex_mtdt_path)) hex_mtdt_path
)

compile_pti_data(
  shp_path       = "app-data/shapes.rds",
  metadata_paths = metadata_paths,
  output_dir     = "app-data",
  var_overrides  = var_overrides
)
```

## Data summary

A bird's-eye view of every compiled indicator -- pillar, inclusion
flags, source admin level, and the value distribution. Use it to
confirm the compile did what you expected before launching.

```{r}
pti_summary_table(
  fct_template_reader("app-data/metadata.xlsx"),
  type = "overview"
)
```

## Preview locally

Edit `landing-page.md` (in the project root) to describe your app,
then launch it:

```{r}
#| eval: false
launch_pti(
  shp_dta  = readRDS("app-data/shapes.rds"),
  inp_dta  = fct_template_reader("app-data/metadata.xlsx"),
  app_name = "My country PTI"
)
```
